Charged pion emission from neutron-rich heavy ion collisions for studies on the symmetry energy

### 2020-07-18 chapters/ch_data_analysis.tex
Submitted to Committee

>>>>

We can write the Gaussian fit for the pions as,

\begin{equation}
G_\pi(z) = A_\pi e^{-\frac{(x-\mu_\pi)^2}{2\sigma_\pi^2}},
\end{equation}

and for the background spectra, in the pion cases $e^+$ and $e^-$:

\begin{equation}
G_B(z) = A_B e^{-\frac{(x-\mu_\pi)^2}{2\sigma_\pi^2} }.
\end{equation}

The probability that a particle is a pion can be described as,

\begin{equation}
P(\pi) = \left( 1 + \frac{A_B}{A_\pi} e^{-\frac{(x-\mu_\pi)^2}{2\sigma_\pi^2}  + \frac{(x-\mu_B)^2}{2\sigma_B^2} } \right)^{-1}
\label{eq:pionProbs}
\end{equation}

The probability value will be used later as a weighting factor when filling the histogram track-by-track as will be discussed in Section~\ref{sec:corrSpectra}.

### 2020-07-18 chapters/ch_data_analysis_2.tex
Submitted to Committee

\begin{equation}
C_a = \frac{4\pi}{\Delta\Omega},
\end{equation}

@@

\subsection{Correcting Pion Spectra (Efficiency and Acceptance)}

To correct for the efficiency, we apply a track-by-track correction, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the efficiency database, from the parameters discussed in Section~\ref{sec:efficiency}. The correction factor $C_i$ is defined as,

\begin{equation}
C_i = \epsilon_i^{-1}.
\end{equation}

@@

>>>>

\begin{equation}
C_a = \frac{4\pi}{\Delta\Omega},
\label{eq:acceptCorrFactor}
\end{equation}

@@

\subsection{Correcting Pion Spectra (Efficiency and Acceptance)}
\label{sec:corrSpectra}
To correct for the efficiency, we apply a track-by-track correction, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the efficiency database, from the parameters discussed in Section~\ref{sec:efficiency}. The correction factor $C_i$ is defined as,

\begin{equation}
C_i = \epsilon_i^{-1}.
\label{eq:effCorrFactor}
\end{equation}

@@

\subsection{Pion Statistical Error}

The treatment of statistical error that will be discussed here applies to the total pion yield and the binned contents of the pion spectral ratio. The total pion multiplicity is defined as the number of pions per event. Accounting for all the corrections and cuts discussed above this can be expressed as,

\begin{equation}
Y(\pi^\pm) = A\sum_{i} \frac{P(\pi)}{\epsilon_i},
\end{equation}

where $P(\pi)$ is the probability that a track is a pion described in Eq.~\ref{eq:pionProbs}, $\epsilon_i(p_{Lab}, \theta_{Lab})$ is the efficiency of particle $i$ given by its momentum $p_{Lab}$ and polar angle $\theta_{Lab}$ as discussed in Section~\ref{sec:efficiency}. The factor $A$ is a scale factor defined as,

\begin{equation}
 A = \frac{C_a}{N_{total}},
\end{equation}

where $N_{total}$ is the total number of events, $C_a$ is the correction factor accounting for the solid angle measured as given in Eq.~\ref{eq:acceptCorrFactor}. The statistical uncertainty is broken into two different parts, due to the way ROOT calculates the error of each histogram bin entry with the Sumw2() function, which calculates the error of each bin as,

\begin{equation}
{\delta b}^2 = \sum_i^{N_{\pi}} {\delta w_i}^2,
\end{equation}

where the weight of the bin is $w_i = A P(\pi) \epsilon_i^{-1}$. This still does not account for the efficiency error which was calculated using the MC embedding method. The efficiency class returns the error of each efficiency bin which scales like the total number of input track $\sqrt{N_{MC}}$. The extra error coming from the efficiency error can be expressed as,

\begin{equation}
{\delta \epsilon }^2 = A^2 \sum_i^{N_{\pi}} \left[ \left( \frac{P(\pi)^{\pi} }{\epsilon_i} \right)^2  \left(  \frac{P(\pi)^{\pi} }{\epsilon_i} \frac{\delta \epsilon_i}{\epsilon_i} \right)^2   \right]
\end{equation}

These two errors make up the total statistical error accounting for the measured number of particles, via Poisson statistics, and the efficiency error following error propagation. They are combined in quadrature such that the total error is $\delta \sigma^2 = \delta\epsilon^2 + \delta b^2$. The systematic error analysis is discussed in Appendix~\ref{sec:cutvar}. This discussion focuses on the systematics originating from the particular cuts we take in our analysis. As will be seen in the results section, the pion ratios seem to cancel out any of these effects and therefore any systematics introduced through our cuts need not be added into the total error bars. For the single particle yields some systematic error bars are added as guided by the cut variation analysis.

### 2020-07-18 chapters/ch_results.tex
Submitted to Committee

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{ypt_sn132.png}
\caption{Rapidity and transverse momentum plot for $\pi^-$ and $\pi^+$ in the $\tin{132}{124}$ system.}
\label{fig:ptrap_sn132}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{ypt_sn108.png}
\caption{Rapidity and transverse momentum plot for $\pi^-$ and $\pi^+$ in the $\tin{108}{112}$ system.}
\label{fig:ptrap_sn108}
\end{figure}

The integrated pion yield for both systems, and the $\pi^-/\pi^+$ ratio, is listed in Table~\ref{tb:pionyield}, where the systematic errors are the first error bar and the statistical error is listed next. This represents data with an estimated average impact parameter of $b_{avg} = \SI{2.2}{\femto\metre}$ and $b_{max} < \SI{3}{\femto\metre}$. Notice that the pion ratio is significantly greater than the $N/Z$ of the system.  This is not completely unexpected; in the delta resonance model mentioned in Section~\ref{sec:pionObs}, we expect  $\pi^-/\pi^+ ~ (N/Z)^2$ \cite{baoan_piprod1,baoan_piprod2}, where N/Z is the ratio of neutron to protons in the dense region where pions are produced. For the $\tin{132}{124}$ system $(N/Z)^2$ = 2.56 and for $\tin{108}{112}$ $(N/Z)^2$ = 1.44, but this n\"aive model assumes no migration of nucleons in or out of the high density region where the isovector mean field and other transport properties may a impact that approximation significantly.

>>>>

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{ypt_sn132.png}
\caption{Rapidity and transverse momentum plot for $\pi^-$ and $\pi^+$ in the $\tin{132}{124}$ system. The measured half of the distribution describes the forward rapidities $y_o > 0$ and has no other extrapolations. Where as the reflected side shows the measured distribution assuming symmetry about $y_o = 0$.}
\label{fig:ptrap_sn132}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{ypt_sn108.png}
\caption{Rapidity and transverse momentum plot for $\pi^-$ and $\pi^+$ in the $\tin{108}{112}$ system. The measured half of the distribution describes the forward rapidities $y_o > 0$ and has no other extrapolations. Where as the reflected side shows the measured distribution assuming symmetry about $y_o = 0$.}
\label{fig:ptrap_sn108}
\end{figure}

The integrated pion yield for both systems, and the $\pi^-/\pi^+$ ratio, is listed in Table~\ref{tb:pionyield}, where the systematic errors are the first error bar and the statistical error is listed next. This represents data with an estimated average impact parameter of $b_{avg} = \SI{2.2}{\femto\metre}$ and $b_{max} < \SI{3}{\femto\metre}$. Notice that the pion ratio is significantly greater than the $N/Z$ of the system.  This is not completely unexpected; in the delta resonance model mentioned in Section~\ref{sec:pionObs}, we expect  $\pi^-/\pi^+ \sim (N/Z)^2$ \cite{baoan_piprod1,baoan_piprod2}, where N/Z is the ratio of neutron to protons in the dense region where pions are produced. For the $\tin{132}{124}$ system $(N/Z)^2$ = 2.56 and for $\tin{108}{112}$ $(N/Z)^2$ = 1.44, but this n\"aive model assumes no migration of nucleons in or out of the high density region where the isovector mean field and other transport properties may a impact that approximation significantly.
